Analysing Constraint Grammars with a SAT-solver

### 2016-03-08 paper/lrec/eval_futurework.tex
more on finnish

\begin{table}[]
\centering
\begin{tabular}{|l|l|l|l|}

\hline
                   & \textsc{nld}  & \textsc{spa}lem  & \textsc{spa}nolem \\ \hline
\# rules           & 59            & 279       & 279     \\ \hline
\# readings        & 336           & 3905      & 1735    \\ \hline
\# true positives  & 7             & n         & m    \\ \hline
\# internal/interaction & 6 + 1    & n         & + 25    \\ \hline

\# false positives & 0             & n         & 5    \\ \hline
\# internal/interaction & 0        & n         & 5 +   \\ \hline

\clock{} with amb. 
           classes & ??s           & 5h 42m    &  11m25  \\ \hline
% \clock{} last rule & ??s           & ??s       & ?min ??s    \\ \hline
\clock{} no amb. 
           classes & 10ish sec       & 14m 16s    & ?h ??min    \\ \hline

@@

\begin{itemize}
\item[] 
\texttt{REMOVE Podar  IF (0 Podar) (1 Adv) (2C Inf OR Enc) ;} \\
\texttt{REMOVE Sentar IF (0 Sentar) (1C Inf OR Enc) ;}
\end{itemize}

@@

\hline
              & 1 clitic + lemmas from grammar & 2 clitics + lemmas from grammar & 1 clitic + all readings 
                                                from grammar & all readings from grammar \\ \hline
\# readings   & 5851 (4263 
                 + 1588) & 9494
                       (7906 + 1588) & 6657 (4263 + 2394) & 2394  \\ \hline
\# conflicts  & 214      & 214       & 24                & 24  \\ \hline
\# interaction & 214      & 214       & 24                & 24  \\ \hline

@@

The second two columns are playing with the full set of readings from the grammar. For both of these, the number of reported conflicts was only 24.

@@

\begin{itemize}
\item[] \item[]\texttt{SELECT "oma" + Gen IF .. (0C Nom) ;}
\item[]\texttt{SELECT ADV (-1C AD) (NOT 0 PP) (NOT 1 A) ;}
\end{itemize}

Both of these are internal conflicts, which may not be trivial to see: 
the first rule requires the target to be genitive and unambiguously nominative: 
the problem in the first rule is that the target is required to be  the problem is that the definition of \texttt{PP} includes \texttt{adv} among others. With the sets expanded, this rule becomes \texttt{SELECT adv IF (NOT 0 pp|adv|adp|po|pr) ...}.

@@

In this case, two rules together remove the target of the third, leaving no way for there to be adjective or noun.

@@

However, many of the use cases do not require running the whole
grammar. Testing the interaction between 5--10 rules takes just
seconds in all languages, if the ambiguity class constraints are not included. 
A downside in the ambiguity classes is that generating them takes a long time, 
and while 5-minute overhead may be acceptable when checking the full grammar,
it is hardly so when analysing just a handful of rules.

>>>>

\begin{table}[]
\centering
\begin{tabular}{|p{2.80cm}|l|l|l|}

\hline
                   & \textsc{nld}  & \textsc{spa}lem  & \textsc{spa}nolem \\ \hline
\# rules           & 59            & 279       & 279     \\ \hline
\# readings        & 336           & 3905      & 1735    \\ \hline
\# true positives 
    \small{(internal + interaction)}
                   & 7 (6 + 1)     & n         & m ( + 25)   \\ \hline
\# false positives 
 \small{(internal + interaction)}
                   & 0        & n         & 5 +   \\ \hline

\clock{} with amb. 
           classes & ??s           & 5h 42m    &  ??m ?s  \\ \hline
\clock{} no amb. 
           classes & 10ish sec       & ?m ?s    & 15m 57s    \\ \hline

@@

\begin{itemize}
\item[]\begin{verbatim}
REMOVE Podar  IF (0 Podar) (..) ;
REMOVE Sentar IF (0 Sentar) (..) ;

\end{verbatim}
\end{itemize}

@@

\hline
              & 1 clitic + 
                lemmas from 
                 grammar & 2 clitics 
                           + lemmas from 
                              grammar & 1 clitic +
                                          all readings 
                                          from grammar    & all readings from grammar \\ \hline
\# readings   & 5851 (4263 
                 + 1588) & 9494
                       (7906 + 1588) & 6657 (4263 + 2394) & 2394  \\ \hline
\# conflicts  & 214      & 214       & 22                & 22  \\ \hline
\# internal/interaction & 211 + 3    & 211 + 3           & 19 + 3    & 19 + 3  \\ \hline

@@

The second two columns are playing with the full set of readings from the grammar. For both of these, the number of reported conflicts was only 22.
Given the preliminary nature of the results, we did not do a full analysis of all the 214 reported conflicts.
We analysed all the 22, and

@@

\begin{itemize}
\item[]\begin{verbatim}
"oma" SELECT Gen IF (..) (0C Nom) ;

SELECT Adv IF (NOT 0 PP) (..) ;
\end{verbatim}
\end{itemize}

Both of these are internal conflicts, which may not be trivial to see.
The first rule requires the target to be genitive and unambiguously nominative; however, these two tags cannot combine in the same reading.
As for the second rule, the definition of \texttt{PP} includes \texttt{adv} among others---with the sets expanded, this rule becomes \texttt{SELECT adv IF (NOT 0 pp|adv|adp|po|pr) ...}.

@@

The first two rules together remove the target of the third, leaving no way for there to be adjective or noun.

@@

However, many of the use cases do not require running the whole
grammar. Testing the interaction between 5--10 rules takes just
seconds in all languages, if the ambiguity class constraints are not included. 
A downside in the ambiguity classes is that generating them takes a long time, 
and while the overhead may be acceptable when checking the full grammar,
it is hardly so when analysing just a handful of rules.
